Performance Models for Data Transfers: A Case Study with Molecular Chemistry Kernels

### 2019-04-12 ICPP-19/communications.tex
minor

	Given a schedule, \scomm($i$) and \scomp($i$) represent the
	start times of task $i$ on communication and computation
	resources. A schedule is feasible if for every time $t$, the
	amount of memory required by all tasks such that $\scomm(i) \leq t
	\leq \scomp(i) + COMP_i$ is not more than the memory capacity
	$C$. 
	\todo[inline]{SKumar: Why not a direct statement?  For simplicity, we assume in all of this paper that the amount of memory required by a task is equal to its communication volume.}
	For simplicity, we assume in all of this paper that the
	amount of memory required by a task is proportional to its
	communication time, and without loss of generality we consider
	that memory size of a task is equal to its communication time.

@@

		As \textit{Transfer time = Latency + (Data Volume)/Bandwidth}. For simplicity, we consider input transfer time is same as input data volume (i.e., \textit{Latency = 0} and \textit{Bandwidth = 1}). Otherwise, we can adjust $C$ such that at any point in a schedule at max one $K_i$ and three $A_i$ tasks can be active.

@@

		With these $K_i$ tasks in place, the schedule on the computation
		resource contains $m$ slots of length exactly $b'$, in which all $A_i$
		tasks must fit without preemption. We can thus define triplet $TR_i$
		as the setof tasks which execute during the communication phase of
		task $K_i$. Since at each point in $S$, exactly one $K_i$ task is
		active, and since $S$ has no idle time on the computation resource,
		the total execution time of tasks in $TR_i$ is exactly $b'$, and thus
		$a_{i_1} + a_{i_2} + a_{i_3} = b$. This partition is thus a valid
		solution for the \threepart instance $A$.

@@

	\subsection{Dynamic Selection}
	Dynamic strategies are based on the following principle: when the communication resource is idle, a task is chosen based on a ranking criterion which differ depending on the heuristic, among those which fit in memory. For example, if the selection criteria is to choose a highly compute intensive task, then we compute the ratio of computation time and communication time for all tasks, and we select a task with the maximum ratio among which induce minimum idle time on the computation resource and fit in the currently available memory. If no task fits in memory then we leave the resource idle at that point and proceed to the next event point. We also ensure that the order on both resources is the same, by ordering tasks on the computation resource accordingly.

@@

	In all strategies, the initial order is obtained by Algorithm~\ref{alg:OrderOfExecutionInfinteMemory}. We define the following heuristics based on how we select a task from the set of tasks which fit in memory and induces minimum idle time on the computation resource. If no task fits in memory then we leave the resource idle and forward to the next event point.
	%%	\todo[inline]{LED: What happens if no task fit in memory?}

@@

		\subsection{Experimental Settings}
		\label{sec:expSetting}

@@

		Figures ~\ref{fig:ratio_to_optimal_hf} and ~\ref{fig:ratio_to_optimal_ccsd} show the distribution of the performance of each heuristic for all considered memory capacities, where plots are categorized by memory capacities. For each memory capacity and each heuristic, the box on the plot displays the median, first and last quartile, and the whiskers indicate minimum and maximum values, with outliers are shown by black dots.
		\subsection{HF Performance}	
		\begin{figure*}[htb]
			\includegraphics[scale=0.5]{./all-binpack/ratio_to_optimal_selected_hf.pdf}
			\caption{Comparison of different heuristics for HF with $m_c=..$.}
			\label{fig:ratio_to_optimal_hf}
		\end{figure*}

		As indicated above, HF tasks operate on less heterogeneous tiles, this is also noticeable in Figure~\ref{fig:ratio_to_optimal_hf}. All heuristics depict similar behavior for minimum memory capacity $m_c$ and increasing the memory capacity slightly does not change the performance of all heuristics. As memory capacity is increased further, dynamic variants of heuristics start performing better. For the moderate memory capacities (close to $2m_c$), static order with dynamic correction variants outperform others. $GG$ heuristic does not achieve good performance, because its task sequence is obtained considering no extra memory is available, but is then applied in a different scenario where memory is limited. Surprisingly, the $BP$ heuristic which considers only memory constraint obtains good performance for a static heuristic, but is outperformed by more dynamic approaches.
		%%		\todo{you mean infinite memory? SKumar: GG is classic in no-wait (or no buffer) 2-macine flowshop}

@@

		Figure~\ref{fig:ratio_to_optimal_best_hf} shows the performance comparison of the best variant in each category, in addition to the \textit{order of submission} ($OS$) strategy which orders tasks in the (arbitrary) order in which they are given. Static strategies are expected to perform better when there is not any memory capacity restriction, and indeed this plot shows that static strategies face capacity bottleneck and underperform with limited memory. Dynamic strategies achieve slightly better performance with limited memory capacity, but when memory capacity is larger, static order with dynamic correction strategies outperform all others.

@@

		\subsection{Scheduling in Batches}
		In most applications, the runtime scheduler may only observe a limited batch of independent tasks. Therefore we organize tasks of each trace file in batches of 100 (the last batch may have less than 100 tasks). We apply each heuristic on each group independently. Figure~\ref{fig:best_variants_batch} shows the performance of the best variants of each category for both applications. The plots exhibit behavior similar to Figures~\ref{fig:ratio_to_optimal_best_hf} and~\ref{fig:ratio_to_optimal_best_ccsd}: static order with dynamic corrections variants attain maximum communication-computation overlap and outperform other heuristics.

>>>>

	Given a schedule, \scomm($i$) and \scomp($i$) represent the
	start times of task $i$ on communication and computation
	resources. A schedule is feasible if for every time $t$, the
	amount of memory required by all tasks such that $\scomm(i) \leq t
	\leq \scomp(i) + COMP_i$ is not more than the memory capacity
	$C$. 
	\todo[inline]{SKumar: How about this?  For simplicity, we assume throughout the paper that tasks do not have any intermediate memory requirement. Hence, amount of memory required by a task is equal to its communication volume.}
	\todo[inline]{Skumar2LED: Actually adding internal memory requirement will make NP-completeness proof straightforward. I started with this problem ;).}
	For simplicity, we assume in all of this paper that the
	amount of memory required by a task is proportional to its
	communication time, and without loss of generality we consider
	that memory size of a task is equal to its communication time.

@@

		\todo[inline]{SKumar: Do we need to mention here separately that memory requirement is equal to communication volume? I think no. right?}
		As \textit{Transfer time = Latency + (Comm Volume)/Bandwidth}. For simplicity, we consider  transfer time is same as communication volume (i.e., \textit{Latency = 0} and \textit{Bandwidth = 1}). Otherwise, we can adjust $C$ such that at any point in a schedule at max one $K_i$ and three $A_i$ tasks can be active.

@@

		With these $K_i$ tasks in place, the schedule on the computation
		resource contains $m$ slots of length exactly $b'$, in which all $A_i$
		tasks must fit without preemption. We can thus define triplet $TR_i$
		as the set of tasks which execute during the communication phase of
		task $K_i$. Since at each point in $S$, exactly one $K_i$ task is
		active, and since $S$ has no idle time on the computation resource,
		the total execution time of tasks in $TR_i$ is exactly $b'$, and thus
		$a_{i_1} + a_{i_2} + a_{i_3} = b$. This partition is thus a valid
		solution for the \threepart instance $A$.

@@

	\subsection{Dynamic Selection}
	Dynamic strategies are based on the following principle: when the communication resource is idle, a task is chosen based on a selection criterion which differ depending on the heuristic, among those which fit in memory and induce minimum idle time on the computation resource.  For example, if the selection criterion is to choose a highly compute intensive task, then we compute the ratio of computation time and communication time for all tasks, and we select a task with the maximum ratio among those which induce minimum idle time on the computation resource and fit in the currently available memory. If no task fits in memory then we leave the resource idle at that point and proceed to the next event point. We also ensure that the order on both resources is the same, by ordering tasks on the computation resource accordingly.

@@

	In all strategies, the initial order is $OMIM$ order,  obtained by Algorithm~\ref{alg:OrderOfExecutionInfinteMemory}. We define the following heuristics based on how we select a task from the set of tasks which fit in memory and induce minimum idle time on the computation resource. If no task fits in memory then we leave the resource idle and forward to the next event point.
	%%	\todo[inline]{LED: What happens if no task fit in memory?}

@@

		\section{Experimental Settings}
		\label{sec:expSetting}

@@

		Figures ~\ref{fig:ratio_to_optimal_hf} and ~\ref{fig:ratio_to_optimal_ccsd} show the distribution of the performance of each heuristic for different memory capacities, where plots are categorized by memory capacities. For each memory capacity and each heuristic, the box on the plot displays the median, first and last quartile, and the whiskers indicate minimum and maximum values, with outliers are shown by black dots.
		\subsection{HF Performance}	
		\begin{figure*}[htb]
			\includegraphics[scale=0.5]{./all-binpack/ratio_to_optimal_selected_hf.pdf}
			\caption{Comparison of different heuristics for HF with $m_c=..$.}
			\label{fig:ratio_to_optimal_hf}
		\end{figure*}

		As indicated above, HF tasks operate on less heterogeneous tiles, this is also noticeable in Figure~\ref{fig:ratio_to_optimal_hf}. All heuristics depict similar behavior for minimum memory capacity $m_c$ and increasing the memory capacity slightly does not change the performance of all heuristics. As memory capacity is increased further, dynamic variants of heuristics start performing better. For the moderate memory capacities (close to $2m_c$), static order with dynamic corrections variants outperform others. $GG$ heuristic does not achieve good performance, because its task sequence is obtained considering no extra memory is available, but is then applied in a different scenario where memory is limited. Surprisingly, the $BP$ heuristic which considers only memory constraint obtains good performance for a static heuristic, but is outperformed by more dynamic approaches.
		%%		\todo{you mean infinite memory? SKumar: GG is classic in no-wait (or no buffer) 2-macine flowshop}

@@

		Figure~\ref{fig:ratio_to_optimal_best_hf} shows the performance comparison of the best variant in each category, in addition to the \textit{order of submission} ($OS$) strategy which orders tasks in the (arbitrary) sequence in which they are given. Static strategies are expected to perform better when there is not any memory capacity restriction, and indeed this plot shows that static strategies face capacity bottleneck and underperform with limited memory. Dynamic strategies achieve slightly better performance with limited memory capacity, but when memory capacity is larger, static order with dynamic correction strategies outperform all others.

@@

		\subsection{Scheduling in Batches}
		In most applications, the runtime scheduler may only observe a limited batch of independent tasks. Therefore we organize tasks of each trace file in batches of 100 (the last batch may have less than 100 tasks). We apply each heuristic on each group in succession. Figure~\ref{fig:best_variants_batch} shows the performance of the best variants of each category for both applications. The plots exhibit behavior similar to Figures~\ref{fig:ratio_to_optimal_best_hf} and~\ref{fig:ratio_to_optimal_best_ccsd}: static order with dynamic corrections variants attain maximum communication-computation overlap and outperform other heuristics.
